---
title: "Run MFDP with the Classic Input Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{run-classic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

R MFDP supports running forecast using the classic Lowestoft's MFDP
input data. A single input data consists of several files including
its index file and a control file.
The `mfdp` package provides samples of these input files
[here](https://github.com/REDUS-IMR/mfdp/tree/main/inst/sample/nea-had-2020).

## Files structure

This set of input files is indexed by an index file. Let's have a peek at the content
in one of the index file (e.g., `Fpa2020ind.txt`):

**Fpa2020ind.txt**
```
Fpa2020MFDP Index file 27.04.2020
 0 
****
Fpa2020CWt.txt
Fpa2020SWt.txt
Fpa2020M.txt
Fpa2020Mat.txt
Fpa2020PF.txt
Fpa2020PM.txt
Fpa2020F.txt
Fpa2020N.txt
Fpa2020Ctrl.txt
```

The above index file points to several other files. Let's have a look inside 
the control file (`Fpa2020Ctrl.txt`):

**Fpa2020Ctrl.txt**
```
HCR2020MFDP Index file 27.04.2020control file27.04.2020
5
4,7
497416,293623,38636,272681,272681,
1
0
0
215000
1.248296
.929582
```

And one of the data file (`Fpa2020CWt.txt`)

**Fpa2020CWt.txt**
```
Fpa2020CWt
1,3
2020,2024
3,13
1
.716,.95,1.224,1.604,1.895,2.077,2.38,2.626,2.79,3.045,3.329,
.708,.946,1.192,1.471,1.846,2.128,2.311,2.627,2.806,2.993,3.417,
.744,.957,1.182,1.43,1.736,2.087,2.358,2.573,2.807,3.004,3.376,
.744,.957,1.182,1.43,1.736,2.087,2.358,2.573,2.807,3.004,3.376,
.744,.957,1.182,1.43,1.736,2.087,2.358,2.573,2.807,3.004,3.376,
```

You can find the full description of the files in the official
Lowestoft's MFDP manual document [link](https://www.researchgate.net/publication/296788449_The_Multi-Fleet_Deterministic_Projection_Program_MFDP).

## Running the forecast

Running the forecast using the classic input is straightforward:

```{r}
# Load the mfdp library
library(mfdp)

# Get the index file (note that we need the full path)
input <- system.file("sample/nea-had-2020", "Fpa2020ind.txt", package = "mfdp")
print(input)

# Run it
output <- mfdp(input, run_name = "test", out_dir = "out")
```

## Output Files

Running the `mfdp` function above will instruct MFDP to run the forecast
and write output files into `./out` directory. We can check the output files:

```{r}
list.files("out")
```

## Results

After successfully running a forecast, MFDP returns several output
objects in a single list. Here we use the term **prm** for the 
**Management Options Table** and **prs** for the 
**Single Option Prediction** results.

```{r}
print(names(output))

# Let's see the Management Options Table
print(output[["prm"]])
```

The **stk** object is an `FLStock` class object <https://flr-project.org/FLCore/reference/FLStock.html>
from the `FLR` suite <https://flr-project.org/>. The developers have provide a wealth of tutorial about the
format here: <https://flr-project.org/#tutorials>.

Next we have the **fmult** object, which holds the *F multiplier* values over the years in the `FLQuant` class,
the **ftgt** object which is the *F target* over the years, and lastly **parameters** object, which is
the control parameters/configuration for the forecast.

Some example of methods supported by FLStock:

```{r}
library(FLCore)

# Get the SSB
print(ssb(output[["prs"]]$stk))

# Get the Catch total
print(catch(output[["prs"]]$stk))

# Get the Catch number
print(catch.n(output[["prs"]]$stk))

# Get the Fbar
print(fbar(output[["prs"]]$stk))
```

Let's plot the forecast results:

```{r}
# Plot the Management Options Table
plot(output[["prm"]]$stk)
```

```{r}
# Single Option Prediction
plot(output[["prs"]]$stk)
```