---
title: "Run MFDP with a Single Input File"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{run-single}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In addition to the classic Lowestoft's MFDP input data,
R MFDP also supports running forecast using a single file for its input data.
This new method simplify the run for the users as they can
have a quick overview of the data quickly.

Please note you will need to provide a control file if you
want to use a single input data file.

The `mfdp` package provides two samples of the single input files:
[here](https://github.com/REDUS-IMR/mfdp/tree/main/inst/sample).
In the same directory you can also find a sample of the control file.

## TLDR

For the impatient, here is how to quickly run a forecast using the provided sample files:

```{r, eval = FALSE}
# Load library
library(mfdp)

# Get input data path
input_file <- system.file("sample", "nea-had-2020.input.csv", package = "mfdp")
# Get param data path
config_file <- system.file("sample", "nea-had-2020.params.txt", package = "mfdp")

# Reading a single file input must be done in two steps
## Parse the input as FLStock object
input <- read_mfdp_input_table(input_file)
## Run the forecast
output <- mfdp(input, configs = config_file, run_name = "test", out_dir = "out")
```

For the more detailed descriptions, please continue below.

## Input data file structure

The structure of the input data file is identical to the Lowestoft's MFDP
`.prd` file output. Here is an example of the data:

**nea-had-2020.input.csv**
```
"MFDP version 1a"
"Run: Manage2020"
"Time and date: 15:02 27.04.2020"
"Fbar age range: 4-7"
""
""
2020
"Age","N","M","Mat","PF","PM","SWt","Sel","CWt"
3,497416,.402,.027839644,0,0,.276890309,.037,.716
4,531880,.269,.084263959,0,0,.518907563,.154,.95
5,171000,.238,.230460922,0,0,.859979101,.324,1.224
6,60331,.273,.526207605,0,0,1.35966736,.504,1.604
7,29429,.2,.779874214,0,0,1.854188211,.524,1.895
8,10839,.2,.884133612,0,0,2.16838843,.584,2.077
9,9746,.2,.963917526,0,0,2.784901758,.584,2.38
10,3557,.2,.976530612,0,0,3.015625,.584,2.626
11,4436,.2,1,0,0,3.542497377,.584,2.79
12,1896,.2,1,0,0,3.361,.584,3.045
13,4665,.2,1,0,0,3.628,.584,3.329
""
2021
"Age","N","M","Mat","PF","PM","SWt","Sel","CWt"
3,293623,.402,.029,0,0,.269,.037,.708
4,".",.269,.083,0,0,.515,.154,.946
5,".",.238,.215,0,0,.823,.324,1.192
6,".",.273,.457,0,0,1.214,.504,1.471
7,".",.2,.756,0,0,1.774,.524,1.846
8,".",.2,.904,0,0,2.285,.584,2.128
9,".",.2,.948,0,0,2.585,.584,2.311
10,".",.2,.985,0,0,3.222,.584,2.627
11,".",.2,1,0,0,3.416,.584,2.806
12,".",.2,1,0,0,3.726,.584,2.993
13,".",.2,1,0,0,3.677,.584,3.417
""
2022
"Age","N","M","Mat","PF","PM","SWt","Sel","CWt"
3,38636,.402,.029,0,0,.303,.037,.744
4,".",.269,.086,0,0,.526,.154,.957
5,".",.238,.21,0,0,.812,.324,1.182
6,".",.273,.436,0,0,1.168,.504,1.43
7,".",.2,.693,0,0,1.589,.524,1.736
8,".",.2,.888,0,0,2.191,.584,2.087
9,".",.2,.959,0,0,2.718,.584,2.358
10,".",.2,.976,0,0,3.001,.584,2.573
11,".",.2,1,0,0,3.624,.584,2.807
12,".",.2,1,0,0,3.596,.584,3.004
13,".",.2,1,0,0,4.029,.584,3.376
""

"Input units are thousands and kg - output in tonnes"
```

## Control file structure

The control file must be a text file. Here is a sample of the
configuration file:

**nea-had-2020.params.txt**
```
* example param file for NEA Haddock 2020
Ftarget 0.35 * target F in HCR
Fpa 0.47 * Fpa
Flim 0.77 * Flim
Fagerange 4 7 * age range for reference F (if not given anywhere else)
Bpa 80000 * trigger point in HCR
3yearrule no * use 3-rule in calculating TAC, yes/no
TACpreviousyear 215000 * TAC in intermediate year, used in connection with calculating TAC in first year
maxchangeup 25 * max annual increase in TAC (%)
maxchangedown 25 * max annual decrease of TAC (%)
```

> NOTE: The keywords are **case sensitive** and all the values
are **mandatory**.

## Running the forecast

Running the forecast using the single input is a two-steps procedure:

### Loading the data

The first step is we load the data into an `FLStock` class object:

```{r}
# Load the mfdp library
library(mfdp)

# Load the input data
input_file <- system.file("sample", "nea-had-2020.input.csv", package = "mfdp")
input <- read_mfdp_input_table(input_file)

print(input)
```

You might want to examine the input data to make sure everyting are properly
loaded:
```{r}
# Load the FLCore library
library(FLCore)

# Do checks
print(catch.n(input))

print(catch.wt(input))

print(stock.n(input))

print(stock.wt(input))

print(ssb(input))

print(fbar(input))

# and so on...

```

### Run forecast using the control file

Now that you are confident that the data is properly loaded, we can continue
to run the forecast:
```{r}
# Get the config file path
config_file <- system.file("sample", "nea-had-2020.params.txt",
                  package = "mfdp")

# Do forecast
output <- mfdp(input, configs = config_file, run_name = "test", out_dir = "out")
```

## Output Files

Running the `mfdp` function above will instruct MFDP to run the forecast
and write output files into `./out` directory. We can check the output files:

```{r}
list.files("out")
```

> NOTE: Since we use a control file that contains **Fpa**, **Flim**, etc. we are
presented with several output files from different F used in the forecast. Note 
also the summary file that collated the results summary from the runs.

## Results

After successfully running a forecast, MFDP returns several output
objects in a single list.

The first level will have the sub-list of **prm** for the 
**Management Options Table** and **prs** for the **Single Option Prediction** results and  
a **summary table**.

```{r}
print(names(output))
```

Let's see the summary from the different F basis:
```{r}
kableExtra::kable_styling(knitr::kable(output$summary),
            bootstrap_options = "striped", latex_options = "striped")
```

Each **prm** and **prs** sub-lists contains the different basis of F used in the forecast.

```{r}
# Use the F management
print(names(output[["prm"]]))

# Let's see the Management Options Table for Fmgmt
print(output[["prm"]][["fmgmt"]])
```

The **stk** object is an `FLStock` class object <https://flr-project.org/FLCore/reference/FLStock.html>
from the `FLR` suite <https://flr-project.org/>. The developers have provide a wealth of tutorial about the
format here: <https://flr-project.org/#tutorials>.

Next we have the **fmult** object, which holds the *F multiplier* values over the years in the `FLQuant` class,
the **ftgt** object which is the *F target* over the years, and lastly **parameters** object, which is
the control parameters/configuration for the forecast.

Some example of methods supported by FLStock:

```{r}
library(FLCore)

# Get the SSB
print(ssb(output[["prm"]][["fmgmt"]]$stk))

# Get the Catch total
print(catch(output[["prm"]][["fmgmt"]]$stk))

# Get the Catch number
print(catch.n(output[["prm"]][["fmgmt"]]$stk))

# Get the Fbar
print(fbar(output[["prm"]][["fmgmt"]]$stk))
```

Let's plot the forecast results:

```{r}
# Plot the Management Options Table
plot(output[["prm"]][["fmgmt"]]$stk)
```

```{r}
# Single Option Prediction
plot(output[["prs"]][["fmgmt"]]$stk)
```